---
title: "Slack"
format: html
embed-resources: true
execute:
  message: false
code-fold: true  
---

## Slack metrics

```{r}
#| include: false
library(tidyverse)
library(janitor)
library(NHSRplotthedots)

slack <- read_csv("20240426-slack.csv") |> 
  janitor::clean_names()

# Dates are character and have 4 digit years, replace with 2 digit years
# Then convert to date format ymd before finding start of month

slack_clean <- slack |> 
  mutate(new_date = gsub( "([0-9]{2})([0-9]{2})$" , "\\2", date),
         new_date = as.Date(new_date, format = "%d/%m/%y"),
         month = lubridate::floor_date(new_date, "month"))

```

Data is extracted from Slack Workspace Analytics and exported from `r min(slack_clean$new_date)` to `r max(slack_clean$new_date)`

## Active users

Whilst there is no standard for measuring community engagement there are three often used:

- Daily Active Users (DAU)
- Weekly Active Users (WAU)
- Monthly Active Users (MAU) 

Slack doesn't provide Monthly Active Users.

Slack defines an Active member as a person who has read or sent a message in at least one channel or a direct message in the workspace.

```{r}
ggplot(slack_clean, aes(x = new_date, y = daily_active_members)) +
  geom_line() +
  geom_smooth() +
  labs(x = "date",
       y = "percentage",
       title = "Daily Active members") 
```

```{r}
ggplot(slack_clean, aes(x = new_date, y = weekly_active_members)) +
  geom_line() +
  geom_smooth() +
  labs(x = "date",
       y = "percentage",
       title = "Monthly Active members") 
```

## Last 6 months

Looking at the chart over 6 months shows consistent dips in activity at the weekend, as well as key holiday periods like Christmas and Easter.

```{r}
last_6_months <- slack_clean |> 
  filter(new_date >= lubridate::today() %m-% months(6)) |> 
  mutate(day_names = lubridate::wday(new_date, label = TRUE),
         is_weekend = ifelse(day_names %in% c("Sat", "Sun"), 0, 1))

  ggplot(last_6_months, aes(x = new_date, y = daily_active_members)) +
  geom_line() +
  geom_smooth() +
  labs(x = "date",
       y = "percentage",
       title = "Daily Active members") 
```  

Removing the weekends and then using {NHSRplotthedots} to create an SPC chart:

```{r}  
  last_6_months |> 
  filter(is_weekend == 1) |> 
  ptd_spc(value_field = daily_active_members,
          date_field = new_date,
          improvement_direction = "increase")
```

## Active posters and readers

Built in Slack analytics show the active people in a workplace split by active members and active posting members.
Active members can mean people in the workspace who read or have posted which is a useful measure as knowledge sharing is a key part of the community.

```{r}
#| label: active-readers

posting_reading <- last_6_months |> 
  mutate(per_posters_readers = daily_members_posting_messages/daily_active_members * 100)

ggplot(posting_reading, aes(x = new_date, y = per_posters_readers)) +
  geom_line() +
  labs(x = "date",
       y = "percentage",
       title = "Percentage of daily members posting by daily active (posting and reading)") +
  geom_smooth()

```

```{r}
posting_reading |> 
  select(new_date,
         per_posters_readers) |> 
  mutate(day_names = lubridate::wday(new_date, label = TRUE),
         is_weekend = ifelse(day_names %in% c("Sat", "Sun"), 0, 1)) |> 
  filter(is_weekend == 1) |> 
  ptd_spc(value_field = per_posters_readers,
          date_field = new_date,
          improvement_direction = "increase")

```

## Insights

A potential risk of concentrating on the high and low peaks of activity can be that they don't necessarily indicate healthy discussion or engagement.
However, a couple of insights are useful for communication plans relating to sharing knowledge:

- Interactions on Slack show a weekly pattern with weekends and public holidays seeing a drop in numbers.
- Distinct numbers of people posting was particularly high during the conference period in November 2023 as people were directed to Slack specifically to ask questions.